---
title: "Untitled"
author: "Samuel Onalaja"
date: "9/18/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries

```{r}
library(tidyverse)
library(corrplot)
library(mlbench)
library(caret)
library(skimr)
library(mice)
library(purrr)
library(ggplot2)
library(ggthemes)
library(cowplot)
library(class)
library(e1071)
library(grid)
library(tidyr)
library(stringr)
library(naniar)
library(car)

```


Read in data and perform some summary statistics on the raw data
```{r}
Life<-read.csv("/Users/olusho/Documents/School/Stats 2 proj 1/Life Expectancy Data.csv",header=T)
summary(Life)
dim(Life)
skim(Life)
head(Life)
str(Life)
attach(Life)

```

Check for missing values and address it
```{r}
vis_miss(Life)
md.pattern(Life)
```

Remove Missig values

```{r}
#Life2 <- Life %>% drop_na(GDP, BMI)
Life2 <- na.omit(Life)
head(Life2)
str(Life2)
skim(Life2)
```

Convert all categorical variables to factor
```{r}
Life22 =Life2
Life22$Status = as.factor(Life2$Status)
```

##Data Exploaration
checking for correlation Numerical variable vs response
- thinness..1.19.years  and thinness.5.9.years are correlated
- Polio, Diphtheria and Hepatitis.B are correlated
- Measles and HIV.AIDS are correlated 
- GDP, percentage.expenditure and population are correlated
- under.five.deaths and infant.deaths are correlated

```{r}
Attr <- "Life.expectancy"

# Name explanatory variable
ExplVar<- Life2%>% keep(is.numeric) %>% colnames

# Create function
PlotFunc <- function(df, explanatory, response) {
  ggplot(data = df) + geom_density(aes_string(x = explanatory), alpha = 0.5) + xlab(explanatory) + ylab("Life.expectancy")
}

  # Density plot
PlotFunc(Life2, explanatory =  "Alcohol", response = "Life.expectancy")

#  Create plot list for plot_grid function to reference
PlotList <- lapply(ExplVar, function(x) PlotFunc(Life2, x, Life.expectancy))

#  Grid of all categorical variables plotted against Attrition
plot_grid(plotlist = PlotList)
```

##Data Exploaration
checking for correlation with VIF
```{r}
#The vif model shows error due alias coefficient
Life.model<-lm(Life.expectancy~.,data=Life2)  
vif(Life.model)[,3]^2

#Check for alias coeffiecient and drop them
ld.vars <- attributes(alias(Life.model)$Complete)$dimnames[[1]]  
ld.vars

#According to ld.vars "StatusDeveloping" is a problem so we are going to drop it
drop <- c("Status")
Life3 <- Life2[, !(names(Life2) %in% drop)]

#The new vif model worked after dropping status
Life.model2<-lm(Life.expectancy~.,data=Life3)  
vif(Life.model2)[,3]^2

#cor(Life2)


# This table shows the correlation between the numerical variables
corrdfTraintable <- Life2 %>% keep(is.numeric) %>% na.omit %>% cor %>% view

Life2 %>% keep(is.numeric) %>% na.omit %>% cor %>% corrplot("upper", addCoef.col = "black", number.digits = 2, number.cex = 0.5, method="shade", order="hclust", tl.srt=45, tl.cex = 0.8)
invisible(view(corrdfTraintable))


```


```{r}
#removes Infant.death,under.five.deaths, thinness_1.9_years, and percentage_expenditure, GDP, Schooling.
life_new<- Life3[-c(5,7,11,16,18,21)]
life_new.model<- life_new 
life_newModel <- lm(Life.expectancy~.,data=life_new.model)  
vif(life_newModel)[,3]^2
summary(life_newModel)


par(mfrow=c(2,2))
plot(life_newModel)

residplot = resid(life_newModel)
qqnorm(residplot)
qqline(residplot)


Attr <- "Life.expectancy"

# Name explanatory variable
ExplVarNew<- life_new.model%>% keep(is.numeric) %>% colnames

# Create function
PlotFuncNew <- function(df, explanatory, response) {
  ggplot(data = df) + geom_density(aes_string(x = explanatory), alpha = 0.5) + xlab(explanatory) + ylab("Life.expectancy")
}

  # Density plot
PlotFuncNew(Life2, explanatory =  "Alcohol", response = "Life.expectancy")

#  Create plot list for plot_grid function to reference
PlotListNew <- lapply(ExplVarNew, function(x) PlotFuncNew(life_new.model, x, Life.expectancy))

#  Grid of all categorical variables plotted against Attrition
plot_grid(plotlist = PlotListNew)


```


```{r backwards}
#life_backward <- step(life_newModel, direction = "backward")
#summary(life_backward)
```


```{r forwards}
#life_forward <- step(life_newModel, direction = "backward")
#summary(life_forward)
```


```{r}
splitPerc = .8

TrainedLife = sample(1:dim(Life3)[1],round(splitPerc * dim(Life3)[1]))
train = Life3[TrainedLife,]
test = Life3[TrainedLife,]

Life_model <- lm(Life.expectancy~.,data=train)  
vif(Life_model)[,3]^2
summary(Life_model)








Predict_Life3 <- predict(Life_model,interval = "predict",newdata = test)

MSPE <- data.frame(Predictions = Predict_Life3[,1], Actual = test$Life.expectancy)
MSPE$Res <- MSPE$Predictions - MSPE$Actual
MSPE$SqRes <- MSPE$Res^2
mean(MSPE$SqRes)


```



```
